---
title: "Week4_Quiz"
author: "Anyi Guo"
date: "16/01/2019"
output: pdf_document
---

# Week 4 Quiz

## Q1 

Load the vowel.train and vowel.test data sets:
```{r}
library(ElemStatLearn)
library(caret)
library(randomForest)
set.seed(33833)
data(vowel.train)
data(vowel.test)

```

Set the variable y to be a factor variable in both the training and test set. Then set the seed to 33833. Fit (1) a random forest predictor relating the factor variable y to the remaining variables and (2) a boosted predictor using the "gbm" method. Fit these both with the train() command in the caret package.

What are the accuracies for the two approaches on the test data set? What is the accuracy among the test set samples where the two methods agree?



```{r}
# factorize y
vowel.train$y<-as.factor(vowel.train$y)
vowel.test$y<-as.factor(vowel.test$y)

# random forest model. accuracy is 0.5974
mod1<-train(y~.,method="rf",data=vowel.train)
pred1<-predict(mod1,vowel.test)
confusionMatrix(vowel.test$y,pred1)

# alternative to caret: random forest with randomForest package. Accuracy is 0.5887
mod3<-randomForest(y~.,data=vowel.train,importance=TRUE)
pred3<-predict(mod3,vowel.test)
confusionMatrix(vowel.test$y,pred3)

# gbm model. Accuracy is 0.5173
mod2<-train(y~.,method="gbm",data=vowel.train)
pred2<-predict(mod2,vowel.test)
confusionMatrix(vowel.test$y,pred2)

# keep only the observations in the test set which the two models' predictions agree.
#Accuracy is 0.6474 for model 1 (random forest), about 0.6442 for model 2(gbm)

d<-vowel.test[pred1==pred2,]
pred4<-predict(mod1,d)
confusionMatrix(d$y,pred4)
pred5<-predict(mod2,d)
confusionMatrix(d$y,pred5)
```

## Answer to Q1: 

* RF Accuracy = 0.6082

* GBM Accuracy = 0.5152

* Agreement Accuracy = 0.6361



## Q2
Load the Alzheimer's data using the following commands
```{r}
library(caret)

library(gbm)

set.seed(3433)

library(AppliedPredictiveModeling)

data(AlzheimerDisease)

adData = data.frame(diagnosis,predictors)

inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]

training = adData[ inTrain,]

testing = adData[-inTrain,]
```

Set the seed to 62433 and predict diagnosis with all the other variables using a random forest ("rf"), boosted trees ("gbm") and linear discriminant analysis ("lda") model. Stack the predictions together using random forests ("rf"). What is the resulting accuracy on the test set? Is it better or worse than each of the individual predictions?

```{r}
set.seed(62433)

# model 1: random forest
mod1<-train(diagnosis~.,method="rf",data=training)
# model 2: boosted trees
mod2<-train(diagnosis~.,method="gbm",data=training)
# model 3: linear discriminant analysis
mod3<-train(diagnosis~.,method="lda",data=training)


pred1<-predict(mod1,testing)
pred2<-predict(mod2,testing)
pred3<-predict(mod3,testing)

# Model 1 accuracy: 0.7805 
confusionMatrix(testing$diagnosis,pred1)
# Model 2 accuracy: 0.8293  
confusionMatrix(testing$diagnosis,pred2)
# Model 3 accuracy: 0.7683
confusionMatrix(testing$diagnosis,pred3)

# stacking the 3 models together. Accuracy for the stacked model is: 0.8293
predDF<-data.frame(pred1,pred2,pred3,y=testing$diagnosis)
combModFit<-train(y~.,method="rf",data=predDF)
pred4<-predict(combModFit,predDF)
confusionMatrix(predDF$y,pred4)
```

## Answer to Q2:
Stacked Accuracy: 0.80 is better than random forests and lda and the same as boosting.